---
title: "20yearCleaningCode"
author: "Matilda Andersson"
date: "2024-06-19"
output: html_document
---

Code to clean and reorganize the fish data to make it usable in a shiny app

orig: Catch_raw_20232019 new: Catch_raw_20232002
orig: Catch_20232002 new: Catch_20232002

load packages
```{r}
library(tidyverse)
```

load data
```{r}
Catch_raw_20232002 <- read_csv("data/Detaljuttag_då_längd_inkl_tomfångst_20year.csv", col_types = cols(
  År1 = col_double(),   # Specify 'species' as character
  Lokal = col_character(),   # Specify as character
  Fångstområde = col_character(),
  Djupstratum1 = col_character(),
  störning1 = col_character(),
  Artbestämning1 = col_character(),
  Längdgrupp = col_double(),
  Station = col_double(),
  Antal1 = col_double(),
  Vikt_uppmätt = col_character(),
  Lat_grader = col_character(), # Specify as character - then remove comma and convert to number
  Long_grader = col_character(), # Specify as character - then remove comma and convert to number
))

EnglishNames <- read_csv("data/FishNamesEnglish_20year.csv")

```

selected required columns and change from , to . as decimal point
```{r}
Catch_20232002 <- Catch_raw_20232002 %>% select(År1, Lokal, Fångstområde, Djupstratum1, störning1, Artbestämning1, Längdgrupp, Antal1, Vikt_uppmätt, Lat_grader, Long_grader, Station)

clean_up_commas <- function(x) {
  x <- gsub(",", ".", x)  # Replace commas with periods
}

Catch_20232002 <- Catch_20232002 %>% 
  mutate(
    Vikt_uppmätt = clean_up_commas(Vikt_uppmätt),
    Djupstratum1 = clean_up_commas(Djupstratum1),
    Lat_grader = clean_up_commas(Lat_grader),
    Long_grader = clean_up_commas(Long_grader))

Catch_20232002 <- Catch_20232002 %>%
  mutate(
    Lat_grader = as.numeric(Lat_grader),
    Long_grader = as.numeric(Long_grader))
```


År1 - Fishing year
Lokal - Region (More general location)
Fångstområde - Location (More specific location)
Djupstratum1 - Net depth
störning1 - Was the net disturbed (only want no)
Artbestämning1 - Species
Längdgrupp - Length (cm)
Antal1 - Number of fish
Vikt_uppmätt - Weight (kg)
Lat_grader - Lat
Long_grader - Long


re-name things 
```{r}
#rename columns
Catch_20232002 <- Catch_20232002 %>% 
  rename(Year = År1,
         Region = Lokal,
         Location = Fångstområde,
         Net_depth = Djupstratum1,
         Disturbance = störning1,
         Species = Artbestämning1,
         Length_cm = Längdgrupp,
         Number = Antal1,
         Weight_kg =  Vikt_uppmätt,
         Latitude = Lat_grader,
         Longitude = Long_grader)
```

group by location -- calculate mean lat and lon by location --> new df
replace existing lat and lon with newly calculated
```{r}
mean_coords <- Catch_20232002 %>%
  group_by(Location) %>%
  summarise(
    avg_lat = mean(Latitude),
    avg_long = mean(Longitude))

Catch_20232002 <- Catch_20232002 %>%
  left_join(mean_coords, by = "Location") %>%
  select(-Latitude, -Longitude) %>%
  rename(Latitude = avg_lat, Longitude = avg_long)

rm(mean_coords)
```

How many species are there - how abundant - decide what to remove
```{r}
Catch_count <- Catch_20232002 %>% 
  count(Species)
write_csv(Catch_count, "Catch_count_20232002.csv")
```

Remove the species that are likely a mistake or were caught only a few times, 
```{r}
RemoveSpecies <- read_csv("data/RemoveSpecies.csv")
RemoveSpecies <- RemoveSpecies %>% pull(Species)

Catch_20232002 <- Catch_20232002 %>%
  filter(!Species %in% c(RemoveSpecies))
```

re-name species -- maybe -- but we can come back to this
```{r}
#rename species to english names
Catch_20232002 <- Catch_20232002 %>%
  left_join(EnglishNames, by = "Species") %>%
  mutate(Species = Species_Eng) %>%
  select(-Species_Eng)
```

Remove any nets that were disturbed -- data is not reliable 
```{r}
Catch_20232002 <- Catch_20232002 %>% 
  filter(Disturbance != 'JA')

#drop the column "disturbance" -- no longer needed
Catch_20232002 <- Catch_20232002 %>% select(!Disturbance)
```

Remove missing data
```{r}
#where length or weight wasn't recorded
Catch_20232002 <- Catch_20232002 %>%
  filter(!is.na(Length_cm) & !is.na(Weight_kg))

# where net wasn't recorded
Catch_20232002 <- Catch_20232002 %>% 
  filter(Net_depth != '-')
```

create one row per individual -- required for the calculations
```{r}
Catch_20232002 <- Catch_20232002 %>%
  uncount(Number)
```

write to .csv for further analysis
```{r}
write.csv(Catch_20232002, "Catch_20232002.csv", row.names = FALSE)
```

