---
title: "20YearCalculateIndicators"
author: "Matilda Andersson"
date: "2024-06-19"
output: html_document
---

load packages
```{r}
library(tidyverse)
```

load data
```{r}
Catch_20232002 <- read_csv("Catch_20232002.csv")
```

calculate effort (number of nets)
You could add net depth if you want to be able to look at CPUE at diff depths or remove one depth from the analysis
```{r}
Effort_Nnets <- Catch_20232002 %>%
  group_by(Year, Location) %>%
  summarise(Effort_Nnets = n_distinct(Station), .groups = 'drop')
```

remove individuals <12 cm -- not caught representativly in the monitoring nets
```{r}
Catch_20232002 <- Catch_20232002 %>%
  filter(Length_cm >= 12)
```

group-by year:location:species -- count individuals (there is one row per individual)
```{r}
Indiv_count <- Catch_20232002 %>%
  group_by(Year, Location, Species) %>%
  summarise(count = n(),
    Latitude = first(Latitude),
    Longitude = first(Longitude),
    .groups = 'drop')
```

calculate CPUE
```{r}
#join the count and effort dataframes
CPUE_data <- Indiv_count %>%
  left_join(Effort_Nnets, by = c("Year", "Location"))

CPUE_results <- CPUE_data %>%
  mutate(CPUE = count / Effort_Nnets)

CPUE_results <- CPUE_results %>%
  filter(!is.na(Species))
```

calculate L90 https://pub.epsilon.slu.se/32950/1/ostman-o-et-al-20240215.pdf
if statements 
if perch -- length >= 15
if flounder -- length >= 15
if whitefish -- legnth >= 25
if zander -- length >= 25
if pike -- length >= 25
if other use the general filter of >= 12
if # of individuals < 50 do not calculate -- error message -- "Your numbers look a bit fishy -- locations have been removed due to low sample size" (Orjan suggests 300 -- but for this project we are using 50)
```{r}
filter(
    (Species == "Perch" & Length_cm >= 15) |
    (Species == "European flounder" & Length_cm >= 15) |
    (Species == "Whitefish" & Length_cm >= 25) |
    (Species == "Zander" & Length_cm >= 25) |
    (Species == "Pike" & Length_cm >= 25) |
    (!Species %in% c("Perch", "European flounder", "Whitefish", "Zander", "Pike") & Length_cm >= 12))

L90 <- L90_data %>%   
  group_by(Year, Location, Species) %>%
  summarise(L90_cm=quantile(Length_cm, probs=0.9),
            Ntot=length(Length_cm))

L90 <- L90 %>%
  filter((Ntot >= 50))
```

merge results and create .csv
```{r}
 Results_20232002<- CPUE_results %>%
  left_join(L90, by = c("Year", "Location", "Species"))

#create a new column with the number of avaliable years, and what years
Results_20232002_years <- Results_20232002 %>%
  group_by(Location, Species) %>%
  summarise(
    nYears = n_distinct(Year),
    Years = paste(unique(Year), collapse = ", "))

Results_20232002 <- Results_20232002 %>%
  left_join(Results_20232002_years %>% select(Location, Species, nYears, Years), by = c("Location", "Species"))

write.csv(Results_20232002, "Results_20232002.csv", row.names = FALSE)
```


